---
title: "Code club project: Assesing whether intra and inter-genomic variation hinder utility of ASVs"
date: "2022-12-11"
output:
 github_document:
  html_preview: false
---

Developed over a series of *code club* episodes led by Pat Schloss to answer an important question in microbiology and develop comfort using tools to develop reproducible data analysis.


### Questions

-   Within a genome, how many distinct sequences of the 16S rRNA gene were present relative to the number of copies per genome? Ho far aprt are these seuences from each other? How does this scale from a genome to kingdoms?
-   Within a taxa (any level), how many ASVs from that taxa are shared with sister taxa? How does this cahnge with taxonomic level? Variable region?


```{r library_calls, include=FALSE}
library(tidyverse)
library(data.table)
library(rmarkdown)
```

### Dependencies:

-   [mothur v.1.48.0](https://github.com/mothur/mothur/releases/tag/v1.48.0)
-   `wget`
-   `r R.version.string`
    -   `tidyverse` (v. `r packageVersion("tidyverse")`)
    -   `data.table` (v. `r packageVersion("data.table")`)
    *   `rmarkdown` (v. `r packageVersion("rmarkdown")`)

### My computer

```{r sesion_info, echo=FALSE}
sessionInfo()
```
