---
title: "Analyzing the sensitivity and specifity of ASVs for descriminating between genomes"
date: "2022-12-11"
output: 
    github_document:
        html_preview: false
editor_options: 
  chunk_output_type: console
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
```


### Need to determine the number of *rrn* operons across genomes

Our analysis will use full length sequnces

```{r read_v19, message=FALSE}
count_tibble <- read_tsv(here("data/processed/rrnDB.count_tibble"),
                         col_types = "cccd")
```

We want to count and plot the number of copies per genome

```{r rrn}
count_tibble %>% 
    filter(region == "v19") %>%
    group_by(genome) %>%
    summarize(n_rrn = sum(count), .groups = "drop") %>%
    ggplot(aes(x=n_rrn)) + geom_histogram(binwidth = 1)

count_tibble %>% 
    filter(region == "v19") %>%
    group_by(genome) %>%
    summarize(n_rrn = sum(count)) %>%
    count(n_rrn) %>%
    mutate(fraction = n / sum(n))
```

We see that most genomes actually have more than one copy of the *rrn* operon. I wonder whether those different copies are the same sequence / ASV..


### Determine number of ASVs per genoem

Considering most genomes have multiple copies of the *rrn* operon, we need to know whether the all have the same ASVs. Otherwise we run the risk of splitting a single genome into multiple ASVs.

```{r}
count_tibble %>%
    group_by(region, genome) %>%
    summarize(n_asv = n(), n_rrn = sum(count), .groups = "drop") %>%
    group_by(region, n_rrn) %>%
    summarize(med_n_asv = median(n_asv),
              mean_n_asv = mean(n_asv),
              lq_n_asv = quantile(n_asv, prob=0.25),
              uq_n_asv = quantile(n_asv, prob=0.75)) %>%
    filter(n_rrn == 7)

count_tibble %>%
    group_by(region, genome) %>%
    summarize(n_asv = n(), n_rrn = sum(count), .groups = "drop") %>%
    ggplot(aes(x=n_rrn, y=n_asv, color=region)) + 
    geom_smooth(method = "lm")
```

Suprisingly (or not!) the number of ASVs increases at a rate of about 2 ASVs per 3 copies of *rrn* operon in the genome. The sub region of the 16S rRNA region have few ASVs per *rrn* operon.


### Determine whether an ASV is unique to genomes they are found in

Instead of looking at the number of ASVs per genome, we want to see the number of genomes per ASV

```{r}
count_tibble %>%
    group_by(region, asv) %>%
    summarize(n_genomes = n()) %>%
    count(n_genomes) %>%
    mutate(fraction = n/sum(n)) %>%
    filter(n_genomes == 1)
```

We see that with full length sequences, thAT 82% of the ASVs were unique to a genome. For the sub regions, about 76% were unique to a genome.



### To be determined...
* Can we correct for over-representation?
* Consider analysis at species, genus, family, etc. levels
* Consider looking at more broad definition of an ASV